Matching Bullets

### 2015-12-15 imaging-paper.Rnw
Remove last few comments, rebuild
typo
references included in the conclusions

\hh{XXX what would be good is to get the line identifications and project them back onto the surface. }

@@

bm$prediction <- c("Tree", "Random Forest")[as.numeric(bm$variable)]
qplot(data=bm, value, reorder(bullet, value/n), size=I(2.5), 
      shape=prediction, colour=prediction, alpha=I(0.6)) + 
  theme_bw()  + ylab("") + 
#  scale_shape_manual(values=c(1,2)) +
  scale_x_continuous("Number of correctly predicted land-to-land matches", 
                     breaks = 3*0:4, limits=c(0,12),
                     minor_breaks = 1:11) +
  theme(legend.position = "bottom")
@
\caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree (maximum possible number is 12). Both results for the tree and the random forest are shown. For each bullet at least three lands are correctly matched. Unknown bullet B is the hardest for the algorithm to match.}
\end{figure}
%
However, this does not carry a lot of weight because of the few bullet comparisons in question: fifteen unknown bullets are successfully matched to two pairs of ten bullets. Matching bullets can only be tested realistically in a much bigger test case. 
Another thing to note about the  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared. Basing this decision on 0.5 is not necessary. In practice, examiners are allowed a third option of `inconclusive'. On a probability spectrum of outcomes we can therefore introduce an interval of `inconclusive' results in the middle of the spectrum. Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree ($x$-axis) and the random forest ($y$-axis). Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only four cases result in a different decision.

@@

The feasibility of a database of ballistic images was evaluated in a 2008 report~\cite{nap:2008} by the National Research Council. The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS. The bottomline of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found. The remaining problem is simply that statistically the quality of the matching algorithm could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system. 
The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing extremely fine marks. 
The report also found, that results from 2d images are improved upon by matches based on 3d images. This is consistent with the importance of features found here: out of the top five features in Table~\ref{tab:importance}, only the number of matches and mismatches are available for a match based on 2d features.

>>>>

@@

bm$prediction <- c("Tree", "Random Forest")[as.numeric(bm$variable)]
qplot(data=bm, value, reorder(bullet, value/n), size=I(2.5), 
      shape=prediction, colour=prediction, alpha=I(0.6)) + 
  theme_bw()  + ylab("") + 
#  scale_shape_manual(values=c(1,2)) +
  scale_x_continuous("Number of correctly predicted land-to-land matches", 
                     breaks = 3*0:4, limits=c(0,12),
                     minor_breaks = 1:11) +
  theme(legend.position = "bottom")
@
\caption{\label{fig:treeresult}Overall predictions for each bullet based on the decision tree (maximum possible number is 12). Both results for the tree and the random forest are shown. For each bullet at least four lands are correctly matched. Unknown bullet B is the hardest for the algorithm to match.}
\end{figure}
%
However, this does not carry a lot of weight because of the few bullet comparisons in question: fifteen unknown bullets are successfully matched to two pairs of ten bullets. Matching bullets can only be tested realistically in a much bigger test case. 
Another thing to note about the  the random forest's error rates is that they are based on probability cutoffs of 0.5, i.e.\ whenever the predicted probability of a match exceeds 0.5, a match is declared. Basing this decision on 0.5 is not necessary. In practice, examiners are allowed a third option of `inconclusive'. On a probability spectrum of outcomes we can therefore introduce an interval of `inconclusive' results in the middle of the spectrum. Figure~\ref{fig:tree-forest} shows a comparison of the predicted probabilities of a match by the tree ($x$-axis) and the random forest ($y$-axis). Overall, there is a lot of agreement; out of 10,800 land-to-land comparisons, only four cases result in a different decision.

@@

The feasibility of a database of ballistic images was evaluated in a 2008 report~\cite{nap:2008} by the National Research Council. The evaluation was investigating the scalability of NIBIN (National Integrated Ballistic Information Network), which uses proprietary matching algorithms provided by IBIS. The bottomline of the report was that in spite of the many technical and practical hurdles, solutions to all but one problem could be found. The remaining problem is simply that statistically the quality of the matching algorithm could not withstand a hugely increased number of records while still maintaining a reasonable workload for forensic examiners, who have to examine possible matches suggested by the system. 
The findings of the NRC report on ballistic imaging are based on two-dimensional greyscale images, which the committee argued were not reliable enough for distinguishing extremely fine marks. This finding coincides with the assessment by de Kinder et al \cite{dekinder:2004} based on the IBIS Heritage system. A further re-assessment in 2015 by de Ceuster et al \cite{deceuster:2014} came to the same conclusions based on the EvoFinder system. 
The NRC report also found, that results from 2d images are improved upon by matches based on 3d images. This is consistent with the importance of features found here: out of the top five features in Table~\ref{tab:importance}, only the number of matches and mismatches are available for a match based on 2d features.
